### 2019-03-15 docs/notebooks/results.Rmd
update results.Rmd

### Visual search tasks for investigating attention
In spite of the many attentional mechanisms proposed for neural networks, and the many parallels and links between studies of vision in psychology, neuroscience, and artificial intelligence, little work has been done to relate the tasks and models used to investigate visual attention to neural networks. We are aware of only one study [@poderCapacityLimitationsVisual2017], which led us to the results that we report below as we replicated it. Like that study, we measured the performance of an exist CNN on a visual search task. In addition, we test whether our results hold for other CNN architectures, and we apply analyses from signal detection theory-based models of attention to understand our results. Our ultimate goal is to provide new methods for comparing neural network output with human behavior and of benchmarking proposed attentional mechansims. First we describe visual search tasks, and then we describe different types of cognitive models used to explain results obtained with these tasks.
We begin with background on visual search tasks. In the "classical" form of the visual search paradigm [@wolfeVisualSearch1998], experimenters present a subject with a search display on which there are discrete items, i.e. objects. The subject is asked to report whether a target is present among distractors. The experimenter sets the probability that the target will be present on any given trial. Experimenters measure the reaction time and/or accuracy of the subjects. Often, the results are presented as a plot of reaction time as a function of set size (the number of discrete items in stimulus). By fitting a line to the plot of reaction times as a function of set size and taking the slope of that line, experimenters produced estimates of processing speed, i.e. "seconds per item". Slopes are almost horizontal when subjects are shown visual search stimuli where targets differ from distractors in some clearly salient **feature**, e.g. a red target among green distractors, and slopes become much steeper when targets differ from distractors in some harder-to-distinguish **spatial conjunction** of features, e.g., a digital 2 among digital 5s. Some examples of these stimuli are shown below, generated with https://github.com/NickleDave/searchstims.

@@

### A psychophysics analysis of visual search shows set size affects remain even when controlling for how discriminable targets are from distractors
The Feature Integration Theory of visual search attributed set size effects to **serial** processing required to bind different types of features [@treismanFeatureintegrationAttention1980]. This conclusion made sense, based on the result that reaction time increased when discriminating targets from distractors required subjects to consider more than one feature, e.g. both color and orientation. However, later work showed that reaction time effects disappear as targets and distractors become more easily discriminable along one feature dimension [@duncanVisualSearchStimulus1989].  
Later researchers, using methods from psychophysics and signal-detection theory, carefully controlled for the discriminability of targets and distractors [@palmerPsychophysicsVisualSearch2000]. This work provided important alternative explanations for the results seen in visual search experiments: attentional effects can be explained in terms of completely parallel but noisy representations that are used by a decision rule [@ecksteinLowerVisualSearch1998; @ecksteinSignalDetectionModel2000; @palmerMeasuringEffectAttention1993; @palmerSetsizeEffectsVisual1994]. This approach also explains errors in terms of low thresholds used by decision making criteria.  We adapted these psychophysics methods to see if they offered additional insights into the attention-like effects we saw in CNNs.

#### Psychometric curves
First we generated 'feature' search stimuli--red vertical rectangle v. green vertical rectangle--where the red and green were more or less discriminable. We did so by taking advantage of RGB color space: by adding more red to the green color, and vice versa, the colors of the target and distractor can be moved closer to each other from the two red and green verticies of the triangle that defines RGB space. (When the color has "all red" and "all green", it becomes yellow, so the stimuli we generated moved closer and closer to yellow.) We then used these stimuli to measure the accuracy of the neural nets we trained as the target and distractor became more and more similar.  
<img src="../static/psychometric_function_fits.png" alt="efficient" style="width: 500px;"/>
#### Threshold v. set size
We then used the psychometric curves to find what level of target-distractor discriminability gave rise to 75% accuracy, known as the "difference threshold". We found this difference threshold for each trained network and each set size.  
<img src="../static/threshold_v_setsize.png" alt="efficient" style="width: 600px;"/>  
Here we see that the difference threshold goes up with set size, meaning more discriminable target-distractor pairs are required as set size increases. In signal detection theory terms, this result is consistent with any theory of attention where decisions are made with a low threshold. (For a detailed walkthrough of how we did this, please see [this notebook](./psychometric.ipynb).)

>>>>

### Visual search tasks for investigating attention
In spite of the many attentional mechanisms proposed for neural networks, and the many parallels and links between studies of vision in psychology, neuroscience, and artificial intelligence, little work has been done to relate the tasks and models used to investigate visual attention to neural networks. Here we study the whether visual search tasks used to investigate attention and the different models of visual attention in the brain provide alternate ways of understanding existing CNNs, comparing their output with human behavior, and benchmarking new attentional mechansims.
We begin with background on visual search tasks. In the "classical" form of the visual search paradigm [@wolfeVisualSearch1998], experimenters present a subject with a search display on which there are discrete items, i.e. objects. The subject is asked to report whether a target is present among distractors. The experimenter sets the probability that the target will be present on any given trial. Experimenters measure the reaction time and/or accuracy of the subjects. Often, the results are presented as a plot of reaction time as a function of set size (the number of discrete items in stimulus). By fitting a line to the plot of reaction times as a function of set size and taking the slope of that line, experimenters produced estimates of processing speed, i.e. "seconds per item". Slopes are almost horizontal when subjects are shown visual search stimuli where targets differ from distractors in some clearly salient **feature**, e.g. a red target among green distractors, and slopes become much steeper when targets differ from distractors in some harder-to-distinguish **spatial conjunction** of features, e.g., a digital 2 among digital 5s. Some examples of these stimuli are shown below, generated with https://github.com/NickleDave/searchstims.

@@

### A psychophysics analysis of visual search shows set size affects remain even when controlling for how discriminable targets are from distractors
The Feature Integration Theory of visual search attributed set size effects to serial processing (Treisman). This conclusion made sense based on the result that reaction time increased when discriminating targets from distractors required considering more than one feature, e.g. both color and orientation. However, later work showed that reaction time effects disappear as targets and distractors become more easily discriminable along one feature dimension (Duncan paper).  
Later researchers, using methods from psychophysics and signal-detection theory, carefully controlled for the discriminability of targets and distractors (Palmer review). This work provided important alternative explanations for the results seen in visual search experiments: attentional effects can be explained in terms of completely parallel but noisy representations that are used by a decision rule (Palmer, Eckstein). This approach also explains errors in terms of low thresholds used by decision making criteria.  We adapted these psychophysics methods to see if they offered additional insights into the attention-like effects we saw in CNNs.  
#### Psychometric curves
First we generated 'feature' search stimuli--red vertical rectangle v. green vertical rectangle--where the red and green were more or less discriminable. We did so by taking advantage of RGB color space: by adding more red to the green color, and vice versa, the colors of the target and distractor can be moved closer to each other from the two red and green verticies of the triangle that defines RGB space. (When the color has "all red" and "all green", it becomes yellow, so the stimuli we generated moved closer and closer to yellow.) We then used these stimuli to measure the accuracy of the neural nets we trained as the target and distractor became more and more similar.  
<img src="../static/psychometric_function_fits.png" alt="efficient" style="width: 500px;"/>
#### Threshold v. set size
We then used the psychometric curves to find what level of target-distractor discriminability gave rise to 75% accuracy, known as the "difference threshold". We found this difference threshold for each trained network and each set size.  
<img src="../static/threshold_v_setsize.png" alt="efficient" style="width: 600px;"/>  
Here we see that the threshold goes up with set size, meaning more discriminable target-distractor pairs are required as set size increases. In signal detection theory terms, this result is consistent with any theory of attention where decisions are made with a low threshold.

For a detailed walkthrough of how we did this, please see [this notebook](./psychometric.ipynb).
